---
title: "Bayesian_demo"
author: "Yaqian Huang"
date: "2023-09-02"
output: html_document
---

```{r setup, include=FALSE,message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(brms)
library(rstan)
library(bayestestR)
library(bayesplot)
library(tidyverse)

```

## Dataset of voicing of glottal fricatives
### 335 tokens of [h] in 91 languages and 35 tokens of [ɦ] in 15 languages at the word-initial position
### research question: to what extent does voicing intensity differ in voiceless and voiced glottal fricatives at the word-initial position? 

```{r loaddf, message=FALSE, warning=FALSE}
df <- read_csv('df.csv')

head(df)
```

## brms model

```{r fitdf,cache=TRUE,warning=FALSE,message=FALSE}

priors_weak <- c(
  # prior for the Intercept (= the reference level)
  prior(normal(0.2, 1), class = Intercept),
  # prior for the fixed effect coefficient for H
  prior(normal(0.2, .5), class = b, coef = lartypeH),
  # prior for the residual variance for gaussian()
  prior(cauchy(0, 0.1), class = sigma)
)

set.seed(62938)
mod <- brm(formula=norm.soe~dur+lartype+(1|family)+(1|area),data=df,
               family = gaussian(), prior = priors_weak,
               save_pars = save_pars(all = T),
               iter=10000,warmup=2000,
               control = list(adapt_delta = 0.999)) # for higher flexibility & treedepth

```

### Summarize the model fit

```{r summarymod}
summary(mod)
```

Extract samples from fitted model

```{r}
as_draws_df(mod)
```


### Visualize distributions of estimates and effects

```{r plotmod}
plot(mod)

mcmc_plot(mod)+theme_bw(base_size = 18)
```

### Posterior predictive check

```{r ppcheckmod}
# Posterior predictive check:
pp_check(mod, ndraws=100, type='dens_overlay')+theme_bw(base_size = 18)

pp_check(mod, ndraws=11, type='hist')

# Plot posterior predictive distribution of statistical summaries:
pp_check(mod, ndraws = 100, type = "stat", stat = "mean")+theme_bw(base_size = 18)

# Plot prior predictive distribution of statistical summaries:
pp_check(mod, ndraws = 100, type = "stat", stat = "mean",
         prefix = "ppd")+theme_bw(base_size = 18)
```

### Calculate Bayes factor

```{r bffit, cache=TRUE,warning=FALSE,message=FALSE}

priors_weak0 <- c(
  # prior for the Intercept (= the reference level)
  prior(normal(0.2, 1), class = Intercept),
  # prior for the residual variance
  prior(cauchy(0, 0.1), class = sigma)
)

set.seed(62938)

mod0 <- brm(formula=norm.soe~dur+(1|family)+(1|area),data=df,
               family = gaussian(), prior = priors_weak0,
               save_pars = save_pars(all = T),
               iter=10000,warmup=2000,
               control = list(adapt_delta = 0.999)) # for higher flexibility & treedepth

options("mc.cores" = 5)

bayes_factor(
    x1 = mod,
    x2 = mod0,
    silent = TRUE,
    repetitions = 10,
    cores = getOption("mc.cores")
)

```

### Calculate 95% highest density interval for uncertainty

All points within this interval have a higher probability density than points outside the interval. The HDI can be used as Credible Interval (CI) to characterize uncertainty of posterior distributions.

```{r hdi}
# 95% HDI
hdi(mod)
plot(hdi(mod))+theme_bw(base_size = 18)

```

### Calculate probability of direction (pd) for effect existence

Pd can be interpreted as the probability (expressed in percentage) that a parameter (described by its posterior distribution) is strictly positive or negative (whichever is the most probable). It is mathematically defined as the proportion of the posterior distribution that is of the median's sign. Although differently expressed, this index is fairly similar (i.e., is strongly correlated) to the frequentist p-value.

```{r pdirection}

# probability of direction
p_direction(mod)
plot(p_direction(mod))+theme_bw(base_size = 18)

# convert to p-value in corresponding lmer model
pd_to_p(pd = p_direction(mod)$pd[3])

```

### Calculate Region of Practical Equivalence (ROPE) for effect significance

ROPE is a region corresponding to a “null” hypothesis. Rather than concluding that an effect is present when it simply differs from zero, we would conclude that the probability of being outside a specific range (ROPE) that can be considered as “practically no effect” (i.e., a negligible magnitude) is sufficient.

```{r rope}
rope(mod)
plot(rope(mod))+theme_bw(base_size = 18)

```

