---
title: "Bayesian_demo"
author: "Yaqian Huang"
date: "2023-09-02"
output: html_document
---

```{r setup, include=FALSE,message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(brms)
library(rstan)
library(bayestestR)
library(bayesplot)
library(tidyverse)

```

## Dataset of voicing of glottal fricatives
### 335 tokens of /h/ in 91 languages and 35 tokens of /ɦ/ in 15 languages at the word-initial position
### research question: to what extent does voicing intensity differ in voiceless and voiced glottal fricatives at the word-initial position? 

```{r loaddf, message=FALSE, warning=FALSE}
df <- read_csv('df.csv')

head(df)
```

## brms model

```{r fitdf,cache=TRUE,warning=FALSE,message=FALSE}

priors_weak <- c(
  # prior for the Intercept (= the reference level)
  prior(normal(0.2, 1), class = Intercept),
  # prior for the fixed effect coefficient for H
  prior(normal(0.2, .5), class = b, coef = lartypeH),
  # prior for the residual variance for gaussian()
  prior(cauchy(0, 0.1), class = sigma)
)

set.seed(62938)
mod <- brm(formula=norm.soe~dur+lartype+(1|family)+(1|area),data=df,
               family = gaussian(), prior = priors_weak,
               save_pars = save_pars(all = T),
               iter=10000,warmup=2000,
               control = list(adapt_delta = 0.999)) # for higher flexibility & treedepth

```

### Summarize the model fit

```{r summarymod}
summary(mod)
```

Extract samples from fitted model

```{r}
as_draws_df(mod)
```


### Visualize distributions of estimates and effects

```{r plotmod}
plot(mod)

mcmc_plot(mod)+theme_bw(base_size = 18)
```

### Posterior predictive check

```{r ppcheckmod}
# Posterior predictive check:
pp_check(mod, ndraws=100, type='dens_overlay')+theme_bw(base_size = 18)

pp_check(mod, ndraws=11, type='hist')

# Plot posterior predictive distribution of statistical summaries:
pp_check(mod, ndraws = 100, type = "stat", stat = "mean")+theme_bw(base_size = 18)

# Plot prior predictive distribution of statistical summaries:
pp_check(mod, ndraws = 100, type = "stat", stat = "mean",
         prefix = "ppd")+theme_bw(base_size = 18)
```

### Calculate Bayes factor

```{r bffit, cache=TRUE,warning=FALSE,message=FALSE}

priors_weak0 <- c(
  # prior for the Intercept (= the reference level)
  prior(normal(0.2, 1), class = Intercept),
  # prior for the residual variance
  prior(cauchy(0, 0.1), class = sigma)
)

set.seed(62938)

mod0 <- brm(formula=norm.soe~dur+(1|family)+(1|area),data=df,
               family = gaussian(), prior = priors_weak0,
               save_pars = save_pars(all = T),
               iter=10000,warmup=2000,
               control = list(adapt_delta = 0.999)) # for higher flexibility & treedepth

options("mc.cores" = 5)

bayes_factor(
    x1 = mod,
    x2 = mod0,
    silent = TRUE,
    repetitions = 10,
    cores = getOption("mc.cores")
)

```

Calculate 95% highest density interval for uncertainty

```{r hdi}
# 95% HDI
hdi(mod)
plot(hdi(mod))+theme_bw(base_size = 18)

```

Calculate probability of direction for effect existence

```{r pdirection}

# probability of direction
p_direction(mod)
plot(p_direction(mod))+theme_bw(base_size = 18)

# convert to p-value in corresponding lmer model
pd_to_p(pd = p_direction(mod)$pd[3])

```

Calculate Region of Practical Equivalence for effect significance

```{r rope}
rope(mod)
plot(rope(mod))+theme_bw(base_size = 18)

```

