---
title: "Visualization_ggplot"
author: "Yaqian Huang"
date: "2023-09-03"
output: html_document
---

```{r setup, include=FALSE,message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)

```

## Dataset of voicing of glottal consonants (Garellek et al., 2021)

```{r loaddf, message=FALSE, warning=FALSE}
df <- read_csv('df-vis.csv')

df.means <- df %>% group_by(language,Filename,Speaker,label,dur,lartype,syltype,syltypegs,family,area,position)%>% 
  select(norm.soe,praat_tier2_duration,percent_voiceless_laryngeal)%>%
  summarise_if(is.numeric, mean, na.rm = T)

head(df)
head(df.means)
```

## set plot style

```{r, message=FALSE, warning=FALSE}

style <- list(
        theme_bw(base_size = 14),
        theme(
                aspect.ratio     = .8,
                legend.position  = "none",
                strip.background = element_blank(),
                strip.text       = element_text(color = "black", hjust = 0, size = 11),
                panel.border     = element_rect(color = "black", fill = NA),
                axis.text.x = element_text(angle=45, hjust=1)
        )
)

style2 <- list(
        theme_bw(base_size = 14),
        theme(
                aspect.ratio     = .8,
                strip.background = element_blank(),
                strip.text       = element_text(color = "black", hjust = 0, size = 11),
                panel.border     = element_rect(color = "black", fill = NA),
        )
)
        
```


## Categorical ~ 0
### What is the distribution of language family in the dataset? area?

```{r, message=FALSE, warning=FALSE}
# barplot of counts
# highlight sample size when n is small

df %>% count(area)
ggplot(df, aes(x=area, fill=area)) + 
  geom_bar()+style
  
# plot only top 10 language family (with NA)
df %>% count(family) %>% arrange(desc(n)) %>% head(11)

df %>% count(family) %>% arrange(desc(n)) %>% head(11) %>% 
  ggplot(aes(x=family, y=n, fill = family))+
  geom_bar(stat='identity')+style

# proportion
# easier interpretation

df %>% group_by(area)%>%
  summarize(n = n())%>%
  mutate(proportion=prop.table(n))%>%
  ggplot(aes(x=area, y=proportion, fill=area)) + 
  geom_bar(stat = 'identity')+style

df %>% count(family) %>% arrange(desc(n)) %>% head(11) %>%
  mutate(proportion=prop.table(n)) %>% 
  ggplot(aes(x=family, y=proportion, fill = family))+
  geom_bar(stat='identity')+style

```

## Numerical ~ 0
### What is the distribution of norm.soe?

```{r, message=FALSE, warning=FALSE}
# histogram
den <- ggplot(df, aes(x=norm.soe))+
  geom_histogram()+
    ggtitle('histogram')+style2
  
# smoothed density
smo <- ggplot(df, aes(x=norm.soe))+
  geom_density(fill='gray', alpha=0.5)+
    ggtitle('density')+style2

gridExtra::grid.arrange(den, smo, nrow=1)
```

## Numerical ~ Numerical
### What is the relationship between norm.soe and percentage of voicing?

```{r, message=FALSE, warning=FALSE}
# scatterplot
sca <- ggplot(df.means,aes(x=100-percent_voiceless_laryngeal, y=norm.soe, alpha=0.5)) + 
  geom_point(size=0.5, position=position_jitter(0.25))+
  scale_x_continuous('% voicing')+
  style2+theme(legend.position = 'none')

# scatterplot + fitted lm through smooth
scalm <- ggplot(df.means,aes(x=100-percent_voiceless_laryngeal, y=norm.soe, alpha=0.5)) + 
  geom_point(size=0.5, position=position_jitter(0.25))+
  geom_smooth(method = 'lm',na.rm = T)+
  scale_x_continuous('% voicing')+
  style2+theme(legend.position = 'none')


gridExtra::grid.arrange(sca, scalm, nrow=1)

```

## Numerical ~ Categorical
### How does mean norm.soe vary across voiceless and voiced glottal fricatives and breathy vowels?

```{r, message=FALSE, warning=FALSE}
# boxplot
box <- ggplot(df.means%>%filter(syltype %in% c("hV","HV","VhV","VHV","Vh","VH","Cbv","CbvC")), aes(x=lartype, color=lartype, fill=lartype, y=norm.soe))+
  geom_boxplot(alpha=0.5, outlier.alpha = 0.1)+
  ggtitle('Boxplot')+
  style2+theme(legend.position = 'none',aspect.ratio = 1.5)


# violin; large n
vio <- ggplot(df.means%>%filter(syltype %in% c("hV","HV","VhV","VHV","Vh","VH","Cbv","CbvC")), aes(x=lartype, color=lartype, fill=lartype, y=norm.soe))+
  geom_violin(alpha=.5,color='black')+
  ggtitle('Violin')+
  style2+theme(legend.position = 'none',aspect.ratio = 1.5)

# jitter; small n; show all data points
jit <- ggplot(df.means%>%filter(syltype %in% c("hV","HV","VhV","VHV","Vh","VH","Cbv","CbvC")), aes(x=lartype, color=lartype, fill=lartype, y=norm.soe))+
  geom_jitter(size=0.7, alpha=0.5)+
  ggtitle('Jittered')+
  style2+theme(legend.position = 'none',aspect.ratio = 1.5)

# bar plot + error bar
bar <- ggplot(df.means%>%filter(syltype %in% c("hV","HV","VhV","VHV","Vh","VH","Cbv","CbvC")), aes(x=lartype, fill=lartype, y=norm.soe,alpha=0.5))+
  stat_summary(fun.y = mean, 
               geom="bar")+
  stat_summary(fun.data = mean_se,
               geom="errorbar", 
               width=0.5)+
  ggtitle('Barplot')+
  style2+theme(legend.position = 'none',aspect.ratio = 1.5)

gridExtra::grid.arrange(bar, jit, vio, box, nrow=1)
```

## Numerical ~ Numerical + Categorical
### How does norm.soe vary over the time course and across voiceless and voiced glottal fricatives and breathy vowels?

```{r, message=FALSE, warning=FALSE}
# splines
# color coded
df %>% filter(syltype %in% c("hV","HV","VhV","VHV","Vh","VH","Cbv","CbvC"))%>%
        ggplot(aes(prop.time, norm.soe, color = lartype))+
        geom_smooth() + style2

```


## Numerical ~ Categorical + Categorical
### How does mean norm.soe vary across voiceless and voiced glottal fricatives and breathy vowels, and word positions?

```{r, message=FALSE, warning=FALSE}
mean_ci <- function(x){
  m = mean(x,na.rm=T)
  se = sd(x,na.rm=T)/sqrt(length(x))
  ql = qnorm(1-0.025)
  c('y'=m, 'ymin'=m-ql*se, 'ymax'=m+ql*se)
}

# violin + jitter + pointrange

df.means %>% filter(syltype %in% c("hV","HV","VhV","VHV","Vh","VH","Cbv","CbvC"))%>%
  mutate(position = factor(position,levels=c('Initial','Medial','Final')))%>%
ggplot(aes(x=lartype,y=norm.soe ,color=lartype))+
  geom_violin(alpha=0.7,na.rm = T,scale = "width")+
  geom_jitter(height=0,width=0.05,size=.3,aes(color=lartype,fill=lartype))+
  stat_summary(fun.data = mean_ci,
               geom="pointrange", 
               fatten = 1.5,
               position=position_dodge(width=0.9),
               size=2,shape=21,
               color="black",
               fill="white")+
  facet_wrap(~position)+
  style2+theme(legend.position = "none",axis.title.x = element_blank()
  )


```

## Numerical ~ Numerical + 2*Categorial
### How does norm.soe vary over the time course and across voiceless and voiced glottal fricatives and breathy vowels, and word positions?


```{r, message=FALSE, warning=FALSE}
df %>% filter(syltype %in% c("hV","HV","VhV","VHV","Vh","VH","Cbv","CbvC"))%>%
  mutate(position = factor(position,levels=c('Initial','Medial','Final')))%>%
        ggplot(aes(prop.time, norm.soe, color = lartype)) +
        geom_smooth() +
        facet_wrap(~ position, scales = "free_x") + style2

```

